Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses

### 2024-12-10 manuscript_agu.tex
Addressing comments

Unlike current general circulation models (GCMs), the storm-resolving version of ICON does not use convective and cloud parameterization but relies on explicit simulation of convection and clouds on the model grid. Subgrid-scale clouds are not resolved, and the grid cell cloud fraction is always either 0 or 100\%. While this makes the code development simpler without having to rely on uncertain parameterizations, it can miss smaller-scale clouds below the grid resolution. Turbulence and cloud microphysics are still parameterized in this model. To account for the radiative effects of subgrid-scale clouds, a cloud inhomogeneity factor is introduced in the model, which scales down the cloud liquid water for radiative calculations. It ranges from 0.4 at lower tropospheric stability (LTS) of 0 K to 0.8 at 30 K. In addition, vertical mixing is enhanced in unstable and humid lower-tropospheric conditions, which reduces the amount of shallow clouds.

@@

We partitioned our data into two mutually exclusive subsets by cyclonic activity. For this purpose, we used a cyclone tracking algorithm to identify extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis and ICON data. We used the open-source cyclone tracking package CyTRACK \citeA{perez-alarcon2024}. Generally, what constitutes an EC is considered relatively arbitrary due to the very large variability of ECs \citeA{neu2013}. We used the mean sea level pressure field and horizontal wind speed fields as input to the CyTRACK algorithm. The algorithm uses pressure and wind speed thresholds as well as tracking across time steps to identify cyclone centers and radii in each time step. With this information, we could classify geographical areas as either cyclonic or non-cyclonic. Due to a relatively small total area covered by cyclones (as identified by the cyclone center and radius), we chose a circle of a double radius (relative to one identified by CyTRACK) centered at the cyclone center as a cyclonic area for every time step and cyclone. All other areas were identified as non-cyclonic. For identifying cyclones in the observations and the reanalyses, ERA5 pressure and wind fields were used as the input to CyTRACK. This is justified by the fact that the large-scale pressure and wind fields in ERA5 are likely sufficiently close to reality. For identifying cyclones in ICON, its own pressure and wind fields were used as the input to CyTRACK, because the model is free-running, and thus the pressure and wind fields are different from reality. Subsetting by proximity to cyclones is a relatively crude measure because it does not take into account the different sectors of cyclones, which are commonly associated with different weather situations. However, this was a choice made for simplicity of the analysis, given the amount of data.

In addition to the above, we partitioned our data into two mutually exclusive subsets based on LTS, which is derived as the difference between the potential temperature at 700 hPa and the surface. Based on a histogram of LTS in ERA5 and MERRA-2 calculated at all voyage tracks and stations (Fig. \ref{fig:lts}), we determined a statistically-based dividing threshold of 12 K for weak stability (< 12 K) and strong stability (>= 12 K) conditions.

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)} strong stability (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years 2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free running). Cyclonic days are expressed as a fraction of the number of days with cyclonic activity, defined as grid points located within a double radius of any cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

We also analyzed the daily cloud cover (total cloud fraction) distribution. This is a measure of cloudiness, irrespective of height, calculated over the course of a day (UTC). A cloud detected at any height means that the lidar profile was classified as cloudy; otherwise, it was classified as a clear sky. When all profiles in a day are taken together, the cloud cover for the day is defined as the fraction of cloudy profiles in the total number of profiles, expressed in oktas (multiples of 1/8).

>>>>

Unlike current general circulation models (GCMs), the storm-resolving version of ICON does not use convective and cloud parameterization but relies on explicit simulation of convection and clouds on the model grid. Subgrid-scale clouds are not resolved, and the grid cell cloud fraction is always either 0 or 100\%. While this makes the code development simpler without having to rely on uncertain parameterizations, it can miss smaller-scale clouds below the grid resolution. Turbulence and cloud microphysics are still parameterized in this model, and aerosols are taken from a climatology. To account for the radiative effects of subgrid-scale clouds, a cloud inhomogeneity factor is introduced in the model, which scales down the cloud liquid water for radiative calculations. It ranges from 0.4 at lower tropospheric stability (LTS) of 0 K to 0.8 at 30 K. In addition, vertical mixing is enhanced in unstable and humid lower-tropospheric conditions, which reduces the amount of shallow clouds.

@@

We partitioned our data into two mutually exclusive subsets by cyclonic activity. For this purpose, we used a cyclone tracking algorithm to identify extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis and ICON data. We used the open-source cyclone tracking package CyTRACK \citeA{perez-alarcon2024}. Generally, what constitutes an EC is considered relatively arbitrary due to the very large variability of ECs \citeA{neu2013}. We used the mean sea level pressure field and horizontal wind speed fields as input to the CyTRACK algorithm. The algorithm uses pressure and wind speed thresholds as well as tracking across time steps to identify cyclone centers and radii in each time step. With this information, we could classify geographical areas as either cyclonic or non-cyclonic. Due to a relatively small total area covered by cyclones (as identified by the cyclone center and radius), we chose a circle of a double radius (relative to one identified by CyTRACK) centered at the cyclone center as a cyclonic area for every time step and cyclone. All other areas were identified as non-cyclonic. For identifying cyclones in the observations and the reanalyses, ERA5 pressure and wind fields were used as the input to CyTRACK. This is justified by the fact that the large-scale pressure and wind fields in ERA5 are likely sufficiently close to reality. \citeA{mcerlich2023} have shown that wind is simulated well in ERA5 relative to the WindSat polarimetric microwave radiometer measurements \cite{meissner2009}. For identifying cyclones in ICON, its own pressure and wind fields were used as the input to CyTRACK, because the model is free-running, and thus the pressure and wind fields are different from reality. Subsetting by proximity to cyclones is a relatively crude measure because it does not take into account the different sectors of cyclones, which are commonly associated with different weather situations. However, this was a choice made for simplicity of the analysis, given the amount of data.

In addition to the above, we partitioned our data into two mutually exclusive subsets based on LTS, which is derived as the difference between the potential temperature at 700 hPa and the surface. Based on a histogram of LTS in ERA5 and MERRA-2 calculated at all voyage tracks and stations (Fig. \ref{fig:lts}), we determined a statistically-based dividing threshold of 12 K for weak stability ($<$ 12 K) and strong stability ($\geq$ 12 K) conditions.

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)} strong stability (LTS $\geq$ 12 K) time steps in \textbf{(a, c)} ERA5 in years 2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free running). Cyclonic days are expressed as a fraction of the number of days with cyclonic activity, defined as grid points located within a double radius of any cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

We also analyzed the daily cloud cover (total cloud fraction) distribution. This is a measure of cloudiness, irrespective of height, calculated over the course of a day (UTC). A cloud detected at any height means that the lidar profile was classified as cloudy; otherwise, it was classified as a clear sky. When all profiles in a day are taken together, the cloud cover for the day is defined as the fraction of cloudy profiles in the total number of profiles, expressed in oktas (multiples of 1/8). The same calculation is done the lidar observations as for the simulated lidar profiles. We use the term ``okta'' independently of its use in instantaneous synoptic observations, and here it simply means 1/8 (0.125\%) of the daily cloud cover.
